### 2017-08-24 ms/figures_supp.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{lineno}
\linenumbers

\begin{document}
\pagenumbering{gobble}
\beginsupplement

\begin{figure}
\centering
\includegraphics{../figures/env_A1B2.pdf}
\caption{Frequency of phenotype \emph{A} after every three generations in
selection regime $A1B2$. Comparison of dynamics starting with
different initial frequency of phenotype \(A\) (0.01--0.99).
See also Figure 1 and Figure S2. In all cases, $W=1$.}\label{fig:env_A1B2}
\end{figure}

\begin{figure}
\centering
\includegraphics{../figures/AkBk_x0.pdf}
\caption{Convergence of the frequency of phenotype \emph{A} to a stable polymorphism in selection regime \emph{AkBk}.
Comparison of dynamics starting with different initial frequencies of phenotype \(A\) (0.01--0.99), and different $k$, $\rho$ and $w$ values.
The lines show the $x$ frequency of phenotype \emph{A} at the end of each period, after every $2k$ generations.
In all cases, $W=1$.}\label{fig:AkBk_x0}
\end{figure}

\begin{figure}
\centering
\includegraphics{../figures/AkBk_stable_wbar.pdf}
\caption{Stable population mean fitness in selection regime \emph{AkBk} as a function of the vertical transmission rate \(\rho\) and the number \(k\) of generations in which phenotypes \emph{A} and \emph{B} are favored by selection, for different selection intensities.
Colors represent the geometric average of the population mean fitness over \(2k\) generations after \(1,000 \times 2k\) generations of evolution.
Blue markers show the maximum average mean fitness for each period \(k\).
For example, with \(w=0.1\), \(\hat{\rho}=0\) maximizes the average fitness for \(k \le 11\), then \(\hat{\rho}\) increases to \(\hat{\rho} \approx 0.24\), and then continues to decrease as \(k\) increases, down to \(\hat{\rho} \approx 0.15\) for \(k=50\).
Contour lines represent \(\rho\) and \(k\) combinations that produce the same average mean fitness. 
In all cases, $W=1$.}\label{fig:AkBk_stable_wbar}
\end{figure}

\begin{figure}
\centering
\includegraphics{../figures/env_A3B10.pdf}
\caption{Frequency of phenotype \emph{A} after every thirteen generations in
selection regime $A3B10$. Comparison of dynamics starting with
different initial frequency of phenotype \(A\) (0.01--0.99).
In all cases, $W=1$.}\label{fig:env_A3B10}
\end{figure}

\end{document}